---
title: "Algorithm"
bibliography: references.bib
csl: vancouver.csl
vignette: >
  %\VignetteIndexEntry{Algorithm}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
execute: 
  echo: false
---

```{r setup}
#| include: false
library(osdc)
library(dplyr)

algorithm_df <- tibble(name = names(algorithm()), value = algorithm()) |>
  tidyr::unnest_wider(value) |>
  dplyr::select(register, name, title, logic, comments)
```

## General description

The **current implementation** of the algorithm is concisely described
below. A more complete description of the **original implementation** of
the classifier is found in Anders Aasted Isaksen's [PhD
Thesis](https://aastedet.github.io/dissertation/) as well as the
validation paper.[@Isaksen2023]

In brief, this algorithm first classifies those with diabetes from those
without diabetes, using four types of inclusion events. Then, those with
diabetes are further classified as having type 1 diabetes (T1D) based on
patterns in the types of glucose-lowering drugs purchased and in
type-specific diabetes diagnoses recorded. If not classified as type 1
diabetes, the individual is classified as type 2 diabetes (T2D).

The initial **diabetes** classification is defined at the date of the
second occurrence of any of the inclusion events listed below. For
example, an individual with two elevated HbA1c tests followed by a
glucose-lowering drug purchase is included at the latest elevated HbA1c
test. Had the second HbA1c test not been performed (or had it returned a
result below the diagnostic threshold), this person would instead have
been included at the date of the first purchase of glucose-lowering
drugs. Wherever possible, all available data for each event is used,
except for the purchases of glucose-lowering drugs, since the patient
register data on obstetric diagnoses necessary to censor
glucose-lowering drug purchases is only complete from 1997 onwards.

Inclusion criteria are:

1.  HbA1c measurements of ≥48 mmol/mol.
2.  Hospital diagnoses of diabetes.
3.  Diabetes-specific services received at podiatrist.
4.  Purchase of glucose-lowering drugs.

Some of these events are ignored if any of the following exclusion
criteria apply:

1.  HbA1c samples:
    -   Samples taken during pregnancies, as these could potentially be
        due to gestational diabetes mellitus, rather than manifest
        diabetes.
2.  Glucose-lowering drugs:
    -   Purchases of GLP-1 receptor agonists, as this drug class is used
        extensively for weight loss among [individuals without
        diabetes](https://doi.org/10.2147/clep.s456170)
    -   Purchases of some types of SGLT-2 inhibitors recommended in
        Danish guidelines for treatment of [heart
        failure](https://www.cardio.dk/chf#551-farmakologisk-behandling-af-hfref)
        or [kidney
        failure](https://nephrology.dk/vejledninger/ckd-mbd/kronisk-nyresygdom/sglt2i_ckd_uden_dm/)
        among individuals without diabetes
        -   Dapagliflozin & empagliflozin
    -   Purchases made during pregnancies, as these could potentially be
        due to gestational diabetes mellitus, rather than manifest
        diabetes.
    -   Purchases of metformin by women below age 40, as these may be
        treatment for polycystic ovary syndrome (PCOS).
    -   Purchases of metformin with a PCOS-related indication code are
        also ignored.

## High-level flowchart

```{mermaid}
%%| eval: true
%%| fig-cap: "Flow of inclusions and exclusions used for classifying diabetes status using the osdc package. Light blue and orange boxes represent inclusions and exclusions, respectively."

flowchart TD
  subgraph background_population["Background population"]
    lpr_diag[("lpr_diag")]
    lpr_adm[("lpr_adm")]
    kontakter[("kontakter")]
    diagnoser[("diagnoser")]
    sysi[("sysi")]
    sssy[("sssy")]
    lmdb[("lmdb")]
    lab_forsker[("lab_forsker")]
    bef[("bef")]
  end

    background_population --> diabetes_diagnoses["Include:<br>Hospital diagnoses of diabetes"]:::include
    background_population --> hba1c["Include:<br>HbA1c measurements of ≥48 mmol/mol"]:::include

    hba1c --> drop_pregnancies_hba1c["Exclude:<br>HbA1c measurements during pregnancies"]:::exclude

    background_population --> gld["Include:<br>Purchase(s) of glucose-lowering drugs used exclusively for treatment of diabetes"]:::include
    gld --> drop_pcos["Exclude:<br>Metformin purchases in women younger than 40, or with indication code of PCOS"]:::exclude
    drop_pcos --> drop_pregnancies_gld["Exclude:<br>GLD purchases during pregnancies"]:::exclude
    
    background_population --> podiatrist_services["Include:<br>Diabetes-specific podiatrist services"]:::include

    %% Join in classify
    drop_pregnancies_hba1c --> classify_t1d["Classification of type 1 diabetes"]
    drop_pregnancies_gld --> classify_t1d
    diabetes_diagnoses --> classify_t1d
    podiatrist_services --> classify_t1d

    classify_t1d --> classify_t2d["Not type 1 diabetes is classified as type 2 diabetes"]

    background_population ~~~ diabetes_diagnoses
    background_population ~~~ podiatrist_services
    background_population ~~~ hba1c

  %% Styling
  classDef default fill:#EEEEEE, color:#000000, stroke:#000000
  classDef include fill:lightblue
  classDef exclude fill:orange
  style background_population fill:#FFFFFF, color:#000000
```

### Classifying type 1 diabetes

Diabetes type is classified as either T1D or T2D based on patterns of
purchases of insulin drugs (including analogues) and hospital primary
diagnoses of T1D and T2D.

Classification as T1D requires an individual to fulfill either of the
following combinations of criteria:

a.  Must have purchased only insulin drugs and never any other type of
    glucose-lowering drugs, and have at least one primary diagnosis of
    T1D from a medical specialty hospital department.
b.  Must have a majority of T1D diagnoses from endocrinology departments
    (or from other medical specialty departments, in the absence of
    contacts to endocrinology departments), and a purchase of insulin
    within 180 days of their initial glucose-lowering drug purchase,
    with insulin contributing at least two thirds of all defined daily
    doses of glucose-lowering drugs purchased.

### Classifying type 2 diabetes

Individuals not classified as T1D cases are classified as T2D cases.

## Detailed and technical description

Below are the technical, exact implementation of the above description.
These are the logical conditions and exact variables and registers used
that results in the classification of diabetes status and types. They
are shown in the form of tables for each register with a table at the
end showing the logic that is applied across registers.

### `lpr_diag`

```{r}
algorithm_df |> 
  dplyr::filter(register=="lpr_diag") |>
  dplyr::select(-register) |> 
  knitr::kable()
```

### `lpr_adm`

```{r}
algorithm_df |> 
  dplyr::filter(register=="lpr_adm") |>
  dplyr::select(-register) |> 
  knitr::kable()
```

### `diagnoser`

```{r}
algorithm_df |> 
  dplyr::filter(register=="diagnoser") |>
  dplyr::select(-register) |> 
  knitr::kable()
```

### `kontakter`

```{r}
algorithm_df |> 
  dplyr::filter(register=="kontakter") |>
  dplyr::select(-register) |> 
  knitr::kable()
```

### `lab_forsker`

```{r}
algorithm_df |> 
  dplyr::filter(register=="lab_forsker") |>
  dplyr::select(-register) |> 
  knitr::kable()
```

### `lmdb`

```{r}
algorithm_df |> 
  dplyr::filter(register=="lmdb") |>
  dplyr::select(-register) |> 
  knitr::kable()
```

### Across register logic

```{r}
algorithm_df |> 
  dplyr::filter(is.na(register)) |>
  dplyr::select(-register) |> 
  knitr::kable()
```

## References
