---
title: "Algorithm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Algorithm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(osdc)
library(tidyverse)
```

## `lpr_diag`

```{r, echo=FALSE}
algorithm |>
  filter(str_detect(register, "lpr_diag") |>
  knitr::kable()
```

## `lpr_adm`

-   `c_spec` (hospital department) categorize as "endocrinology" if it
    equals 8 or as "other medical" if it is \< 8 or equals either 9 to
    30. 

## `diagnoser`

-   `diagnosekode` starts with "DO0[0-6]", "DO8[0-4]", "DZ3[37]" or
    "DE1[0-4]".
    -   Is T1D if `diagnosekode` starts with "DE10".
    -   Is T2D if `diagnosekode` starts with "DE11".
-   `diagnosetype` is equal to either "A" or "B".
    -   Is a primary diagnosis if it equals "A".
-   `senere_afkraeftet` (if the diagnosis was later retracted) is equal
    to "Nej".

## `kontakter`

-   `hovedspeciale_ans` (hospital department) is categorized as
    "endocrinology" if it equals "medicinsk endokrinologi" or as "other
    medical" if it equals any of "Blandet medicin og kirurgi", "Intern
    medicin", "Geriatri", "Hepatologi", "Hæmatologi",
    "Infektionsmedicin", "Kardiologi", "Medicinsk allergologi",
    "Medicinsk gastroenterologi", "Medicinsk lungesygdomme",
    "Nefrologi", "Reumatologi", "Palliativ medicin", "Akut medicin",
    "Dermato-venerologi", "Neurologi", "Onkologi", "Fysiurgi", or
    "Tropemedicin".

## `lab_forsker`

```{r, echo=FALSE}
algorithm |>
  filter(name == "hba1c") |>
  knitr::kable()
```

## `ssyi` and `sssy`

-   `speciale` starts with "54".
    -   Alternatively, use `spec2` if available.
-   `barnmark` (services provided to a child of the individual) is not
    equal to 0.

## `lmdb`

```{r, echo=FALSE}
algorithm |>
  filter(name == "gld") |>
  knitr::kable()
```

## `bef` and only GLD (via `lmdb`)

To remove those with potential polycystic ovary syndrome:

-   `atc` starts with "A10BA02" and `koen` is equal to 2 (woman) and
    (`date` minus `foed_dato` (birth date) is less than 40 or
    `indication_code` equals one of "0000092", "0000276" or "0000781")
