---
title: "Design"
output: rmarkdown::html_vignette
bibliography: references.bib
csl: vancouver.csl
vignette: >
  %\VignetteIndexEntry{Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(osdc)
```

## Principles

These are the guiding principles for this package:

1.  Functionality is as agnostic to data format as possible (e.g. can be
    used with SQL or Arrow connections, in a data.table format, or as a
    data.frame).
2.  Functions have consistent inputs and outputs (e.g. inputs and
    outputs are the same, regardless of specific conditions).
3.  Functions have predictable outputs based on inputs (e.g. if an input
    is a data frame, the output is a data frame).
4.  Functions have consistent naming based on their action.
5.  Functions have limited additional arguments.
6.  Casing of input variables (upper or lower case) is agnostic, all
    internal variables are lower case, and output variables are lower
    case.

## Use cases

We make these assumptions on how this package will be used, based on our
experiences and expectations for use cases:

-   Entirely used within the Denmark Statistics or the Danish Health
    Authority's servers, since that is where their data are kept.
-   Used by researchers within or affiliated with Danish research
    institutions.
-   Used specifically within a Danish register-based context.

Below is a set of "narratives" or "personas" with associated needs that
this package aims to fulfill:

-   "As a researcher, ..."
    -   "... I want to determine which registers and variables to
        request from Denmark Statistics and Danish Health Data
        Authority, so that I am certain I will be able to classify
        diabetes status of individuals in the registers."
    -   "... I want to easily and simply create a dataset that contains
        data on diabetes status in my population, so that I can begin
        conducting my research that involves persons with diabetes
        without having to tinker with coding the correct algorithm to
        classify them."
    -   "... I want to be informed early and in a clear way whether my
        data fits with the required data type and values, so that I can
        fix and correct these issues without having to do extensive
        debugging of the code and/or data."

## Core functionality

This is the list of functionality we aim to have in the osdc package

1.  Classify individuals type 1 and type 2 diabetes status and create a
    data frame with that information and the date of onset of diabetes.
2.  Provide helper functions to check and process individual registers
    for the variables required to enter into the classifier.
3.  Provide a list of required variables and registers in order to
    calculate diabetes status.
4.  Provide validation helper functions to check that variables match
    what is expected of the algorithm.
5.  Provide a common and easily accessible standard for determining
    diabetes status within the context of research using Danish
    registers.

## Classifier algorithm

A more complete description of the classifier is found in Anders Aasted
Isaksen's [PhD Thesis](https://aastedet.github.io/dissertation/) as well
as the validation paper [@Isaksen2023]. The description below is a brief
and concise version of those documents.

The algorithm for classifying individuals with diabetes is described
below. The overall output of this algorithm is first to classify those
with diabetes, then to further classify and check if the individuals
might have type 1 diabetes, otherwise classify as type 2 diabetes.

Initial **diabetes** classification is defined as the second occurrence
of any of the listed inclusion events. Wherever possible, all available
data for each event is used, except for the purchases of
glucose-lowering drugs, since the data on obstetric diagnoses necessary
to censor glucose-lowering drug purchases is only complete from 1997
onwards. Inclusion criteria are:

1.  HbA1c measurements of ≥48 mmol/mol.
2.  Hospital diagnoses of diabetes.
3.  Diabetes-specific services received at podiatrist.
4.  Purchase of glucose-lowering drugs.

Exclusions are:

1.  HbA1c samples:
    -   Samples taken during pregnancies, as that could be a potential
        gestational diabetes mellitus.
2.  Glucose-lowering drugs:
    -   Brand drugs for weight loss, e.g. *Saxenda*.
    -   Purchases during pregnancies, as that is a potential treatment
        for gestational diabetes mellitus.
    -   Metformin for women below age 40, as that could be a treatment
        for polycystic ovary syndrome.

### Classifying type 1 diabetes

Diabetes type is classified as either T1D or T2D based on patterns of
purchases of insulin drugs (including analogues) and hospital primary
diagnoses of T1D and T2D.

Classification as T1D requires an individual to fulfill either of the
following criteria:

a.  Must have purchased only insulin drugs and never any other type of
    glucose-lowering drugs, and have at least one diagnosis of T1D
b.  Must have a majority of T1D diagnoses from endocrinological
    departments (or from other medical departments, in the absence of
    contacts to endocrinological departments), and a purchase of insulin
    within 180 days after onset of diabetes, with insulin contributing
    at least two thirds of all defined daily doses of glucose-lowering
    drugs purchased.

In populations generated on a fixed index date (such as the
cross-sectional studies associated with this dissertation), individuals
classified as T1D cases must have purchased insulin drugs in the last
year prior to the index date.

### Classifying type 2 diabetes

Individuals not classified as T1D cases are classified as T2D cases.

## Data required from registers

The following is a list of the variables required from specific
registers in order for the package to classify diabetes status:

```{r, echo=FALSE}
variable_description |> 
  mutate(Register = paste0(register_name, "(", register_abbrev, ")")) |> 
  select(Register, Variable = variable_name) |> 
  knitr::kable()
```

## References
