---
title: "Function flow"
output: rmarkdown::html_vignette
bibliography: references.bib
csl: vancouver.csl
vignette: >
  %\VignetteIndexEntry{Function flow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette describes the function conventions and function flow of
the osdc package. The function convention sections go over how we name
functions and how we structure them in terms of input and output. The
function flow describes the functions within the package, both internal
and user-facing, which data sources they rely on, and how they are
connected to each other. First, the functions for classifying diabetes
status are presented, followed by the functions for classifying the
diabetes type.

## Function conventions

The below conventions are *ideals* only, to be used as a guidelines to
help with development and understanding of the code. They are not hard
rules.

### Naming

-   First word is an action verb, later words are objects or conditions.
-   Exclusion criteria are prefixed with `exclude_`.
-   Inclusion criteria are prefixed with `include_`.
-   Helpers that get or extract a condition (e.g., "pregnancy" or "date
    of visit") are prefixed with `get_`.
-   Helpers that drop or keep a specific condition are prefixed with
    `drop_` or `keep_` (e.g., "first visit date to maternal care for
    pregnancy after 40 weeks"). These types of helpers likely are
    contained in the `get_` functions.
-   Helpers that join registers or output of other functions are
    prefixed with `join_`.

### Input and output

-   Few arguments, with one or two core required argument.
-   `include_` functions take a register as the first argument.
    -   One input register database at a time.
-   `exclude_` functions can take a register as the first argument or
    take the output from an `include_` function.
-   Second argument can be an output data from another function.

## Function flow

The osdc package contains one main function that classifies individuals
into those with either type 1 or type 2 diabetes using the Danish
registers: `classify_diabetes()`. This function classifies those with
diabetes (type 1 or 2) based on the Danish registers described in the
`vignette("design")` and `vignette("data-sources")`. All data sources
are used as input for this function. The specific inclusion and
exclusion details are also described in the `vignette("design")`.

This results in the functionality flow for classifying diabetes status
seen below. This flow can be divided into two sections: extracting the
diabetes population and classifying diabetes type which we will detail
in the following sections.

All functions take a `data.frame` type object as input and outputs the
same type of object as the input object (a `data.frame` type). For
instance, if the input is a `data.table` object, the output will also be
a `data.table`.

![Flow of functions, as well as their required input registers, for
classifying diabetes status using the osdc package. Light blue and
orange boxes represent filtering functions (inclusion and exclusion
events, respectively). Uncoloured boxes are helper functions that get or
extract a condition or joins data or function
outputs.](images/function-flow.png)

## Population extraction

In the following sections, we describe the functions used to extract the
diabetes population from the Danish registers. The functions are divided
into inclusion and exclusion events, and the final diagnosis date is
calculated based on these events.

![Flow of functions, as well as their required input registers, for
extracting the population with diabetes using the osdc package. Light
blue and orange boxes represent filtering functions (inclusion and
exclusion events, respectively). Uncoloured boxes are helper functions
that get or extract a condition or joins data or function
outputs.](images/function-flow-population.png)

### Inclusion events

```{r, include=FALSE}
library(dplyr)
library(osdc)
```

#### Hospital diagnoses

#### Joining LPR2 and LPR3 data

The helper functions `join_lpr2()` and `join_lpr3()` join records of
diagnoses to administrative information in LPR2-formatted and
LPR3-formatted data, respectively.

`join_lpr2()` takes `lpr_diag` and `lpr_adm` as inputs, filters to the
necessary diagnoses (`c_diag` starting with "DO0[0-6]", "DO8[0-4]",
"DZ3[37]", "DE1[0-4]", "249", or "250"), joins the required information
by record number (`recnum`), and outputs a `data.frame` with the
following variables:

-   `pnr`: identifier variable
-   `date`: date of the recorded diagnosis (renamed from `d_inddto`)
-   `specialty`: department specialty (renamed from `c_spec`)
-   `diagnosis_code`: diagnosis code (renamed from `c_diag`)
-   `diagnosis_type`: diagnosis type (renamed from `c_diagtype`)

`join_lpr3()` takes `diagnoser` and `kontakter` as inputs, filters to
the necessary diagnoses (`diagnosekode` starting with "DO0[0-6]",
"DO8[0-4]", "DZ3[37]" or "DE1[0-4]"), joins the required information by
record number (`dw_ek_kontakt`), and outputs a `data.frame` with the
following variables:

-   `pnr`: identifier variable (renamed from `cpr`)
-   `date`: date of the recorded diagnosis (renamed from `dato_start`)
-   `specialty`: department specialty (renamed from `hovedspeciale_ans`)
-   `diagnosis_code`: diagnosis code (renamed from `diagnosekode`)
-   `diagnosis_type`: diagnosis type (renamed from `diagnosetype`)
-   `diagnosis_retracted`: if the diagnosis was later retracted (renamed
    from `senere_afkraeftet`)

These outputs are passed to `include_diabetes_diagnoses()` (and to
`get_pregnancy_dates()`, see exclusion events) for further processing
below.

#### Processing of diabetes diagnoses

The function `include_diabetes_diagnoses()` uses the hospital contacts
from LPR2 and LPR3 to include all dates of diabetes diagnoses to use for
inclusion, as well as additional information needed to classify diabetes
type. Diabetes diagnoses from both ICD-8 and ICD-10 are included.

The function takes the outputs of `join_lpr2()` and `join_lpr3()` as
inputs and processes each input separately to generate the following
internal variables:

-   From `join_lpr2`:
    -   `pnr`: identifier variable
    -   `date`: dates of all included diabetes diagnoses:
    -   registered as primary (A) or secondary (B) diagnoses, regardless
        of type or department:
        -   Keep rows where `diagnosis` starts with "DE1[0-4]", "249" or
            "250", and `diagnosis_type` is either "A" or "B"
    -   `is_primary`: Define whether the diagnosis was a primary
        diagnosis (`diagnosis_type` == "A")
    -   `is_t1d`: Define whether the diagnosis was T1D-specific
        (`diagnosis` starts with "DE10" or "249")
    -   `is_t2d`: Define whether the diagnosis was T2D-specific
        (`diagnosis` starts with "DE11" or "250")
    -   `department`: Define whether the diagnosis was made made by an
        endocrinological (if `specialty` == 8 then `department` ==
        "endocrinology") or other medical department (if `specialty` \<
        8 or 9-30 then `department` == "other medical")
-   From `join_lpr3()`:
    -   `pnr`: identifier variable
    -   `date`: dates of all included diabetes diagnoses:
    -   registered as primary (A) or secondary (B) diagnoses, regardless
        of type or department, but exclude retracted diagnoses:
        -   Keep rows where `diagnosis` starts with "DE1[0-4]",
            `diagnosis_type` is either "A" or "B" and
            `diagnosis_retracted` == "Nej"
    -   `is_primary`: Define whether the diagnosis was a primary
        diagnosis (`diagnosis_type` == "A")
    -   `is_t1d`: Define whether the diagnosis was T1D-specific
        (`diagnosis` starts with "DE10")
    -   `is_t2d`: Define whether the diagnosis was T2D-specific
        (`diagnosis` starts with "DE11")
    -   `department`: Define whether the diagnosis was made made by an
        endocrinological department (if `specialty` == "medicinsk
        endokrinologi" then `department` == "endocrinology") or other
        medical department (if `specialty` is any of "Blandet medicin og
        kirurgi", "Intern medicin", "Geriatri", "Hepatologi",
        "Hæmatologi", "Infektionsmedicin", "Kardiologi", "Medicinsk
        allergologi", "Medicinsk gastroenterologi", "Medicinsk
        lungesygdomme", "Nefrologi", "Reumatologi", "Palliativ medicin",
        "Akut medicin", "Dermato-venerologi", "Neurologi", "Onkologi",
        "Fysiurgi", or "Tropemedicin" then `department` == "other
        medical")

Internally, these intermediate results are combined and processed
together. And ultimately, `include_diabetes_diagnoses()` outputs a
single `data.frame` with the following variables (up to two rows per
individual):

-   `pnr`: identifier variable
-   `dates`: dates of the first and second hospital diabetes diagnosis
-   `n_t1d_endocrinology`: number of type 1 diabetes-specific primary
    diagnosis codes from endocrinological departments
-   `n_t2d_endocrinology`: number of type 2 diabetes-specific primary
    diagnosis codes from endocrinological departments
-   `n_t1d_medical`: number of type 1 diabetes-specific primary
    diagnosis codes from medical departments
-   `n_t2d_medical`: number of type 2 diabetes-specific primary
    diagnosis codes from medical departments

This output is passed to the `join_inclusions()` function, where the
`dates` variable is used for the final step of the inclusion process.
The variables of counts of diabetes type-specific primary diagnoses (the four columns prefixed `n_` above) are
carried over for the subsequent classification of diabetes type,
initially as inputs to the `get_t1d_primary_diagnosis()` and
`get_majority_of_t1d_diagnoses()` functions.

#### Diabetes-specific podiatrist services

The function `include_podiatrist_services()` uses `sysi` or `sssy` as
input to extract the dates of all diabetes-specific podiatrist services.

These dates are extracted by filtering values beginning with "54" in the
`speciale` variable of the `sssy` and `sysi` registers by default
(alternatively, the function can take the `spec2` variable as input
instead, if that is the data available to the user). In addition,
services provided to a child of the individual (`barnmak` != 0) are
excluded using the `barnmak` variable. An internal helper function
`get_unique_honuge_dates()` is applied to generate a proper date
variable based on the year-week (wwyy-formatted) variable (`honuge`)
found in the raw data, and de-duplicates multiple services registered on
the same date.

`include_podiatrist_services()` outputs a 2-column data frame with up to
two rows for each individual, containing the following variables:

-   `pnr`: identifier variable
-   `date`: the dates of the first and second diabetes-specific
    podiatrist record

The output is passed to the `join_inclusions()` function for the final
step of the inclusion process.

#### HbA1c tests above the diagnosis cut-off value (48 mmol/mol or 6.5%)

The function `include_hba1c()` uses `lab_forsker` as the input data to
extract the dates of all elevated HbA1c test results, using the
appropriate cut-offs:

-   IFCC units: `analysiscode` NPU27300, any `value` $\geq$ 48 mmol/mol
-   DCCT units: `analysiscode` NPU03835: any `value` $\geq$ 6.5% .

```{r, echo=FALSE}
algorithm |> 
	filter(name=="hba1c") |>
	knitr::kable(caption = "Algorithm used in the implementation for including HbA1c.")
```

Multiple elevated results on the same day within each individual are
deduplicated, to account for the same test result often being reported
twice (one for IFCC, one for DCCT units).

`include_hba1c()` outputs a 2-column data frame containing the following
variables:

-   `pnr`: identifier variable
-   `dates`: the dates of all elevated HbA1c test results

The output is passed to the `exclude_pregnancy()` function for censoring
of elevated results due to potential gestational diabetes (see below).

#### GLD purchases

The function `include_gld_purchases()` uses `lmdb` to extract the dates
of all GLD purchases.

These dates are extracted by including all values beginning with "A10"
in the `atc` variable of the `lmdb` register, except for
glucose-lowering drugs that may be used for other conditions than
diabetes: GLP-RAs (`atc` start with "A10BJ") or
dapagliflozin/empagliflozin (`atc` = "A10BK01" or "A10BK03").

Since the diagnosis code data on pregnancies (see below) is insufficient
to perform censoring prior to 1997, `include_gld_purchases()` only
extracts dates from 1997 onward by default (if Medical Birth Register
data is available to use for censoring, the extraction window can be
extended).

This function outputs a long `data.frame` (since all dates of purchases
must be kept for later use in classifying diabetes type) with the
following variables needed later in the classification part of the
function flow:

-   `pnr`: identifier variable
-   `date`: dates of all purchases of GLD (renamed from `eksd`)
-   `atc`: type of drug
-   `contained_doses`: amount purchased, in number of defined daily
    doses (DDD). Calculated as `volume` (doses contained in the
    purchased package) times `apk` (number of packages purchased)
-   `indication_code`: indication code of the prescription (renamed from
    `indo`)

These events are then passed to a chain of exclusion functions:
`exclude_potential_pcos()` and `exclude_pregnancy()` described in the
sections below.

### Exclusion events

#### Metformin purchases potentially for the treatment of polycystic ovary syndrome

The function `exclude_potential_pcos()` takes the output from
`include_gld_purchases()` and `bef` (information on sex and date of
birth) as inputs and censors (filters out) all purchases of metformin in
women below age 40 at the date of purchase (`atc` = "A10BA02" & `sex` =
"woman" & age at purchase (`date`-`date_of_birth`) \< 40 years) or an
indication code suggesting the prescription was made for treatment of
polycystic ovary syndrome (`atc` = "A10BA02" & `sex` = "woman" &
`indication_code` either of "0000092", "0000276" or "0000781").

This function only performs a filtering operation, and output retains
the same structure and variables as the input passed from
`include_gld_purchases()`. After these exclusions are made, the output
is passed to `exclude_pregnancy()` for further censoring, described
below.

#### HbA1c tests and GLD purchases during pregnancy

The function `exclude_pregnancy()` takes the combined outputs from
`join_lpr2()`, `join_lpr3()`, `include_hba1c()`, and
`exclude_potential_pcos()` and uses diagnoses from LPR2 or LPR3 to
exclude both elevated HbA1c tests and GLD purchases during pregnancy, as
these may be due to gestational diabetes, rather than type 1 or type 2
diabetes.

Internally, this relies on the function `get_pregnancy_dates()` that
uses diagnoses registered in LPR2 and LPR3 to extract
the dates of all recorded pregnancy endings (live births and
miscarriages). These are identified by `diagnosis` values beginning with
"DO0[0-6]", "DO8[0-4]" or "DZ3[37]". The dates output by
`get_pregnancy_dates()` are used to exclude all inclusion events
registered between 40 weeks before and 12 weeks after a pregnancy
ending.

After these exclusion functions have been applied, the output serves as
inputs to two sets of functions:

1.  The censored HbA1c and GLD data are passed to the
    `join_inclusions()` function for the final step of the inclusion
    process.
2.  the censored GLD data is passed to the
    `get_only_insulin_purchases()`,
    `get_insulin_purchases_within_180_days()`, and
    `get_insulin_is_two_thirds_of_gld_doses()` helper functions for the
    classification of diabetes type.

### Join inclusion events

The function `join_inclusions()` appends/row-binds the dates output from
functions the process the four types of inclusion events by `pnr`. Thus,
it takes as input the following variables output from the following
functions:

-   From `include_diabetes_diagnoses()`:
    -   `pnr`: identifier variable
    -   `dates`: dates of the first and second hospital diabetes
        diagnosis
-   From `include_podiatrist_services()`
    -   `pnr`: identifier variable
    -   `dates`: the dates of the first and second diabetes-specific
        podiatrist record
-   From `exclude_pregnancy()`:
    -   `pnr`: identifier variable
    -   `dates`: the dates of the first and second elevated HbA1c test
        results (after censoring)
-   From `exclude_pregnancy()`:
    -   `pnr`: identifier variable
    -   `date`: dates of all purchases of GLD
        -   The dates of the first and second purchase of GLD of each
            individual are extracted from these and appended as two rows
            to the ´dates´ variable.

The output from the function is a `data.frame` containing two variables
(`pnr` and `dates`) and 1 to 8 rows per ´pnr´. This output is passed to
`get_diagnosis_date()`.

### Get diagnosis date

The function `get_diagnosis_date()` takes the output from
`join_inclusions()` and defines the final diagnosis date based on all
the inclusion event types.

First, the inputs are sorted by `dates` within each level of `pnr`, then
the earliest value of `dates` is dropped, so that only those with two or
more events are included. The date of inclusion, `raw_inclusion_date`,
is then defined as the earliest value of `dates`in the remaining rows
for each individual (effectively the date of the second recorded
inclusion event). A third variable, `stable_inclusion_date`, is defined
based on `raw_inclusion_date` (if `raw_inclusion_date` \< stable
inclusion threshold (one year after medication data starts to contribute
to inclusions. Default "31-12-1997"), then `stable_inclusion_date` is
set to `NA`, else it is set to`raw_inclusion_date`). This variable
serves to limit the included cohort to only individuals with valid date
of inclusion (and thereby valid age at inclusion & duration of
diabetes).

`get_diagnosis_date()` outputs a `data.frame` with the following
variables:

-   `pnr`: identifier variable
-   `raw_inclusion_date`: date of inclusion
-   `stable_inclusion_date`: date of inclusion of valid incident cases

This output is passed to the `get_diabetes_type()` function and used to
classify the diabetes type as described below.

### Classifying the diabetes type

The next step of the OSDC algorithm classifies individuals from the
extracted diabetes population as having either T1D or T2D. As described
in the `vignette("design")`, individuals not classified as T1D cases are
classified as T2D cases.

As the diabetes type classification incorporates an evaluation of the
time from diagnosis/inclusion to first subsequent purchase of insulin,
the `get_diabetes_type()` function has to take the date of diagnosis and
all purchases of GLD drugs (after censoring) as inputs. In addition,
information on diabetes type-specific primary diagnoses from hospitals
is also a requirement.

Thus, the function takes the following inputs from
`get_diagnosis_date()`, `exclude_pregnancy()`, and
`include_diabetes_diagnoses()`:

-   From `get_diagnosis_date()`: Information on date of diagnosis of
    diabetes
    -   `pnr`
    -   `raw_inclusion_date`
    -   `stable_inclusion_date`
-   From `exclude_pregnancy()`: Information on historic GLD purchases:
    -   `pnr`: identifier variable
    -   `date`: dates of all purchases of GLD.
    -   `atc`: type of drug
    -   `contained_doses`: defined daily doses of drug contained in
        purchase
-   From `include_diabetes_diagnoses()`: Information on diabetes
    type-specific primary diagnoses from hospitals:
    -   `pnr`: identifier variable
    -   `n_t1d_endocrinology`: number of type 1 diabetes-specific
        primary diagnosis codes from endocrinological departments
    -   `n_t2d_endocrinology`: number of type 2 diabetes-specific
        primary diagnosis codes from endocrinological departments
    -   `n_t1d_medical`: number of type 1 diabetes-specific primary
        diagnosis codes from medical departments
    -   `n_t2d_medical`: number of type 2 diabetes-specific primary
        diagnosis codes from medical departments

For each `pnr` number, several helper functions are applied to these
inputs to extract additional information from the censored GLD data and
diagnoses to use for classification of diabetes type. All of these
return a single value (`TRUE`, otherwise `FALSE`) for each individual:

-   `get_only_insulin_purchases()`:
    -   Inputs passed from `exclude_pregnancy()`:
        -   `atc`
    -   Outputs:
        -   only_insulin_purchases = `TRUE` if no purchases with `atc`
            starting with "A10A" are present
-   `get_insulin_purchases_within_180_days()`
    -   Inputs passed from `exclude_pregnancy()`:
        -   `date` & `atc`
    -   Inputs passed from `get_diagnosis_date()`:
        -   `raw_inclusion_date`
    -   Outputs: `TRUE` If any purchases with `atc` starting with "A10A"
        have a `date` between 0 and 180 days higher than
        `raw_inclusion_date`
-   `get_insulin_is_two_thirds_of_gld_doses()`
    -   Inputs passed from `exclude_pregnancy()`:
        -   `contained_doses` & `atc`
    -   Outputs: `TRUE` If the sum of `contained_doses` of rows of `atc`
        starting with "A10A" (except "A10AE5") is at least twice the sum
        of `contained_doses` of rows of `atc` starting with "A10B" or
        "A10AE5"
-   `get_any_t1d_primary_diagnoses()`:
    -   Inputs passed from `include_diabetes_diagnoses()`:
        -   `n_t1d_endocrinology` & `n_t1d_medical`
    -   Outputs: `TRUE` if the combined sum of the inputs is 1 or above.
-   `get_type_diagnoses_from_endocrinology()`:
    -   Inputs passed from `include_diabetes_diagnoses()`:
        -   `n_t1d_endocrinology`, `n_t2d_endocrinology`
    -   Outputs: `type_diagnoses_from_endocrinology` = `TRUE` if the
        combined sum of the inputs is 1 or above
-   `get_type_diagnosis_majority()`:
    -   Inputs passed from `include_diabetes_diagnoses()`:
        -   `n_t1d_endocrinology`, `n_t2d_endocrinology`,
            `n_t1d_medical` & `n_t2d_medical`
    -   Inputs passed from `get_type_diagnoses_from_endocrinology()`:
        -   `type_diagnoses_from_endocrinology`
    -   Outputs: `TRUE` if `type_diagnoses_from_endocrinology` == `TRUE`
        and `n_t1d_endocrinology` is above `n_t2d_endocrinology`. Also
        `TRUE` if `type_diagnoses_from_endocrinology` = `FALSE` and
        `n_t1d_medical` is above `n_t2d_medical`

`get_diabetes_type()` evaluates all the outputs from the helper
functions to define diabetes type for each individual. Diabetes type is
classified as "T1D" if:

-   `only_insulin_purchases` == `TRUE` & `any_t1d_primary_diagnoses` ==
    `TRUE`
-   Or `only_insulin_purchases` == `FALSE` & `any_t1d_primary_diagnoses`
    == `TRUE` & `type_diagnosis_majority` == `TRUE` &
    `insulin_is_two_thirds_of_gld_doses` == `TRUE` &
    `insulin_purchases_within_180_days` == `TRUE`

`get_diabetes_type()` returns a `data.frame` with one row per `pnr`
number and four columns: `pnr`, `stable_inclusion_date`,
`raw_inclusion_date` & `diabetes_type`. This is the final product of the
OSDC algorithm. See the `vignette("design")` for an more detail on the
two inclusion dates and their intended use-cases.

<!-- TODO: Create updated image similar to https://aastedet.github.io/dissertation/4-results.html#fig-osdc-type-flow to reflect the new diabetes type logic and embed image here for reference-->

<!-- TODO:  The following explanatory sections on T1D and T2D classification need to be aligned with the technical sections above, and possibly moved up to them-->

#### Type 1 classification

The details for the classification of type 1 diabetes is described in
`vignette("design")`. To classify whether an individual has T1D, the
OSDC algorithm includes the following criteria:

1.  `get_t1d_primary_diagnosis()`, which relies on the hospital
    diagnoses extracted from `lpr_diag` (LPR2) and `diagnoser` (LPR3) in
    the previous steps.
2.  `get_only_insulin_purchases()` which relies on the GLD purchases
    from Lægemiddeldatabasen to get patients where all GLD purchases are
    insulin only.
3.  `get_majority_of_t1d_diagnoses()` (as compared to T2D diagnoses)
    which again relies on primary hospital diagnoses from LPR.
4.  `get_insulin_purchase_within_180_days()` which relies on both
    diagnosis from LPR and GLD purchases from Lægemiddelsdatabasen.
5.  `get_insulin_is_two_thirds_of_gld_doses` which relies on the GLD
    purchases from Lægemiddelsdatabasen.

Note the following hierarchy in first function above: First, the
function checks whether the individual has primary diagnoses from
endocrinological specialty. If that's the case for a given person, the
check of whether they have a majority of T1D primary diagnoses are based
on data from endocrinological specialty. If that's not the case, the
check will be based on primary diagnoses from medical specialties.

#### Type 2 classification

As described in the `vignette("design")`, individuals not classified as
type 1 cases are classified as type 2 cases.

## Output

The output of the OSDC algorithm is a `data.frame` which includes four
columns:

1.  **PNR**: The pseudonymised social security number of individuals in
    the diabetes population (one row per individual)
2.  **stable_inclusion_date**: The *stable* inclusion date (i.e., the
    raw date mutated so only individuals included in the time-period
    where data coverage is sufficient to make incident cases
    reliable)[^1]
    <!-- TODO: Specify this time-period: e.g., later than 1997 -->
3.  **raw_inclusion_date**: The *raw* inclusion date (i.e., the date of
    the second inclusion event as described in the [Extracting the
    diabetes population](#extracting-diabetes-population) section above)
4.  **diabetes_type** The classified diabetes type

[^1]: For more information on the "raw" versus "stable" inclusion date,
    see `vignette("design")`.

<!-- TODO: Make sure this is the correct link - and add a link specific to the specific section where this is described -->

For an example, see below.

| PNR        | stable_inclusion_date | raw_inclusion_date | diabetes_type |
|------------|-----------------------|--------------------|---------------|
| 0000000001 | 2020-01-01            | 2020-01-01         | T1D           |
| 0000000004 | NULL                  | 1995-04-19         | T2D           |

: Example rows of the `data.frame` output of the osdc package.

The individuals `0000000001` and `0000000004` have been classified as
having diabetes (`T1D` and `T2D`, respectively). `0000000004` is
classified as having type 1 diabetes (T1D) with an inclusion date of
`2020-01-01`. Since this date is within a time-period of sufficient data
coverage, the column `stable_inclusion_date` is populated with the same
date as `raw_inclusion_date`.

The individual in the second row, `0000000004` is classified as having
type 2 diabetes `T2D` with an inclusion date of `1995-19-04`. Since 1995
is within a time-period of insufficient data coverage,
`stable_inclusion_date` is `NULL`. However, `raw_inclusion_date` still
contains the inclusion date of this individual.

<!-- TODO: Specify the "stable" time-period: e.g., later than 1997 -->

