---
title: "Getting started with osdc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with osdc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package serves two overarching purposes:

1.  To provide an open-source, code-based algorithm to classify type 1
    and type 2 diabetes using Danish registers as data sources.
2.  To inspire discussions within the Danish register-based research
    space on the openness and ease of use on the existing tooling and
    registers, and on the need for an official process for updating or
    contributing to existing data sources.

To read up on the overall design of this package as well as on the
algorithm, check out the `vignette("design")`. For more explanation on
the motivations, rationale, and needs for this algorithm and package,
check out the `vignette("explanation")`.

## Usage

First, let's load up the package, as well as dplyr.

```{r setup}
library(dplyr)
library(osdc)
```

The core of this package depends on the list of variables within
different registers that are needed in order to actually classify the
diabetes status of an individual. This can be found in the dataset:

```{r}
variable_description
```

Let's convert that to a nicer format to show the list as a table.

```{r}
variable_description |> 
  select(register_name, register_abbrev, variable_name, english_description) |> 
  knitr::kable()
```
