---
title: "Getting started with osdc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{osdc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(osdc)
```

When working with the Danish Registers, the choice of the data format is
either SAS or Stata data formats. Which is unfortunate given that both
formats are not well designed to run on massive datasets. So importing
data into R to use osdc is done whichever way you choose, though we do
recommend converting the data files into a Parquet format and using
DuckDB to connect to the files. If you use the R package dplyr, along
with the package arrow, this is very easy to do. We plan on making a
package to help with this process.

<!-- TODO: Other formats you can select? -->

## Steps to classifying diabetes status

```{r}
# 1: Run Import Script to Convert Raw Files from SAS Format --------

# Set path to root of raw files folder, e.g.:
foreign_folder = 'E:/ProjektDB/ALMAU/Rawdata/707517/'

# Set names of files to be imported. Can be split across multiple virtual machines for saving time.
# Executing everything on one machine works fine too.

# To run in 3 simultaneous chuncks on 3 virtual machines, run each of the following sections of 3 lines on a different machine:
sas_files_1 <- "^bef|^lpr_diag|^uaf_diag|^lpr_adm|^uaf_adm|.*sksube201[4-8]"
sas_file_list <- list.files(path = foreign_folder, pattern = sas_files_1, recursive = T)
source(here("R/import.R"))

#
sas_files_2 <- "^sysi|^sssy200[6-9]|^sssy20[1-9]"
sas_file_list <- list.files(path = foreign_folder, pattern = sas_files_2, recursive = T)
source(here("R/import.R"))

#
sas_files_3 <- "^lab|^lmdb"

sas_files_3 <- "^udda|^akm|^faik"
sas_file_list <- list.files(path = foreign_folder, pattern = sas_files_3, recursive = T)
source(here("R/import.R"))

# Set path to root of secondary raw files folder, e.g.:
foreign_folder = 'E:/ProjektDB/ALMAU/Rawdata/707517/Eksterne data/2021-06-09 fra SDS/'
sas_files_3 <- "^lab_forsker"
sas_file_list <- list.files(path = foreign_folder, pattern = sas_files_3, recursive = T)
source(here("R/import.R"))


# Note exclusion of either of sssy and sysi in 2005, where they potentially overlap.
# Note also that we import medication prescriptions from all years, even though we only filter and use those from 1997 onwards.


# 2: Run Filter Script to Create Source Data ---------------------------------

source(here("R/filter.R"))



# 3: Run Population Script to Generate Diabetes Population -------------------
source(here("R/population.R"))


# 4: Draw a Live Sample of the Diabetes Population at a Point in Time--------------------
source(here("R/draw_dm_sample.R"))

# Import the entire diabetes population through time:
dm_population <- readRDS(here('data/output/dm_population.rds'))
# Select which year to draw the population:

dm_2019 <- draw_dm_sample(sample_year = 2019)
saveRDS(dm_2019, here("data/output/dm_population_2019.rds"))

rm(dm_population, dm_2018)

# And draw the corresponding diagrams:
source(here("R/flowcharts.R"))
load(here("data", "source", "flowchart_counts.Rdata"))
draw_inclusion()
draw_type(sample_year = 2019)


# 5: Compress Imported Raw Data to Save Space -----------------------------

compress = TRUE
# Can be set to FALSE if wanting to uncompress the raw data at a later point

source(here("R/compress.R"))
```
