---
title: "Design"
output: rmarkdown::html_vignette
bibliography: ../references.bib
csl: ../vancouver.csl
vignette: >
  %\VignetteIndexEntry{Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(osdc)
```

## Principles

These are the guiding principles for this package:

1.  Functionality is as agnostic to data format as possible (e.g. can be
    used with SQL or Arrow connections, in a data.table format, or as a
    data.frame).
2.  Functions have consistent inputs and outputs (e.g. inputs and
    outputs are the same, regardless of specific conditions).
3.  Functions have predictable outputs based on inputs (e.g. if an input
    is a data frame, the output is a data frame).
4.  Functions have consistent naming based on their action.
5.  Functions have limited additional arguments.
6.  Casing of input variables (upper or lower case) is agnostic, all
    internal variables are lower case, and output variables are lower
    case.

## Use cases

We make these assumptions on how this package will be used, based on our
experiences and expectations for use cases:

-   Entirely used within the Denmark Statistics or the Danish Health
    Authority's servers, since that is where their data are kept.
-   Used by researchers within or affiliated with Danish research
    institutions.
-   Used specifically within a Danish register-based context.

Below is a set of "narratives" or "personas" with associated needs that
this package aims to fulfill:

-   "As a researcher, ..."
    -   "... I want to determine which registers and variables to
        request from Denmark Statistics and Danish Health Data
        Authority, so that I am certain I will be able to classify
        diabetes status of individuals in the registers."
    -   "... I want to easily and simply create a dataset that contains
        data on diabetes status in my population, so that I can begin
        conducting my research that involves persons with diabetes
        without having to tinker with coding the correct algorithm to
        classify them."
    -   "... I want to be informed early and in a clear way whether my
        data fits with the required data type and values, so that I can
        fix and correct these issues without having to do extensive
        debugging of the code and/or data."

## Core functionality

This is the list of functionality we aim to have in the osdc package

1.  Classify individuals type 1 and type 2 diabetes status and create a
    data frame with that information and the date of onset of diabetes.
2.  Provide helper functions to check and process individual registers
    for the variables required to enter into the classifier.
3.  Provide a list of required variables and registers in order to
    calculate diabetes status.
4.  Provide validation helper functions to check that variables match
    what is expected of the algorithm.
5.  Provide a common and easily accessible standard for determining
    diabetes status within the context of research using Danish
    registers.

## Function conventions

Effectively developing both the main user-exposed and internal functions
requires following some conventions and design patterns for building
these functions. There are a few conventions we describe here: naming
patterns for functions and arguments, their argument input requirements,
and their output data structure.

The below conventions are *ideals* only, to be used as a guidelines to
help with development and understanding of the code. They are not hard
rules.

### Naming

-   First word is an action verb, later words are objects or conditions.
-   Exclusion criteria are prefixed with `exclude_`.
-   Inclusion criteria are prefixed with `include_`.
-   Helpers that get or extract a condition (e.g., "pregnancy" or "date
    of visit") are prefixed with `get_`.
-   Helpers that drop or keep a specific condition are prefixed with
    `drop_` or `keep_` (e.g., "first visit date to maternal care for
    pregnancy after 40 weeks"). These types of helpers likely are
    contained in the `get_` functions.
-   Helpers that join the output of other functions are prefixed with
    `join_`.
-   Functions that prepare and process register data are prefixed with
    `prepare_`.

### Input

-   As few arguments as is possible, with as few core required arguments
    as possible (ideally one or two).
-   `include_` functions take a register as the first argument.
    -   One input register database at a time.
-   `exclude_` functions can take a register as the first argument or
    take the output from an `include_` function.
-   All functions take a `data.frame` type object as input. This input
    object doesn't need to be strictly a `data.frame` as long as it acts
    like a `data.frame`. For instance, it could be a `data.table`, a
    `tibble`, or a `duckdb` object.
-   The first argument will always take a data frame type object.
-   The second argument could be an output data frame object from
    another function (usually `include_`).

### Output

-   All functions output the same type of object as the input object (a
    `data.frame` type object). For instance, if the input is a
    `data.table` object, the output will also be a `data.table`.

## Interface

The osdc package contains one main function that classifies individuals
into those with either type 1 or type 2 diabetes using the Danish
registers: `classify_diabetes()`. This function classifies those with
diabetes (type 1 or 2) based on the Danish registers described in the
`vignette("design")` and `vignette("data-sources")`. All data sources
are used as input for this function. The specific inclusion and
exclusion details are described in the `vignette("algorithm")`.

### Input

There is one argument in `classify_diabetes()` for each required
register, so the argument is:

```{r, results='asis'}
registers() |>
  purrr::imap_chr(~ glue::glue("- `{.y}`: The register that is called '{.x$name}' in Danish.")) |>
  unname()
```

### Output

The output is a `data.frame` type object which includes four columns:

-   **pnr**: The pseudonymised social security number of individuals in
    the diabetes population (one row per individual).
-   **stable_inclusion_date**: The *stable* inclusion date (i.e., the
    raw date mutated so only individuals included in the time-period
    where data coverage is sufficient to make incident cases
    reliable)[^1]
    <!-- TODO: Specify this time-period: e.g., later than 1997 -->
-   **raw_inclusion_date**: The *raw* inclusion date (i.e., the date of
    the second inclusion event as described in the [Extracting the
    diabetes population](#extracting-diabetes-population) section above)
-   **diabetes_type** The classified diabetes type

[^1]: For more information on the "raw" versus "stable" inclusion date,
    see `vignette("algorithm")`.

For an example, see below.

| pnr        | stable_inclusion_date | raw_inclusion_date | diabetes_type |
|------------|-----------------------|--------------------|---------------|
| 0000000001 | 2020-01-01            | 2020-01-01         | T1D           |
| 0000000004 | NULL                  | 1995-04-19         | T2D           |

: Example rows of the `data.frame` output of the osdc package.

The individuals `0000000001` and `0000000004` have been classified as
having diabetes (`T1D` and `T2D`, respectively). `0000000001` is
classified as having type 1 diabetes (T1D) with an inclusion date of
`2020-01-01`. Since this date is within a time-period of sufficient data
coverage, the column `stable_inclusion_date` is populated with the same
date as `raw_inclusion_date`.

The individual in the second row, `0000000004` is classified as having
type 2 diabetes `T2D` with an inclusion date of `1995-19-04`. Since 1995
is within a time-period of insufficient data coverage,
`stable_inclusion_date` is `NULL`. However, `raw_inclusion_date` still
contains the inclusion date of this individual.

<!-- TODO: Specify the "stable" time-period: e.g., later than 1997 -->