---
title: "Internal function flow"
---

The function flow describes the functions within the package, both
internal and user-facing, which data sources they rely on, and how they
are connected to each other. First, the functions for classifying
diabetes status are presented, followed by the functions for classifying
the diabetes type.

## Function flow

This results in the functionality flow for classifying diabetes status
seen below. This flow can be divided into two sections: extracting the
diabetes population and classifying diabetes type which we will detail
in the following sections.

![Flow of functions, as well as their required input registers, for
classifying diabetes status using the osdc package. Light blue and
orange boxes represent filtering functions (inclusion and exclusion
events, respectively). Uncoloured boxes are helper functions that get or
extract a condition or joins data or function
outputs.](images/function-flow.svg)

## Population extraction

In the following sections, we describe the functions used to extract the
diabetes population from the Danish registers. The functions are divided
into inclusion and exclusion events, and the final diagnosis date is
calculated based on these events.

![Flow of functions, as well as their required input registers, for
extracting the population with diabetes using the osdc package. Light
blue and orange boxes represent filtering functions (inclusion and
exclusion events, respectively). Uncoloured boxes are helper functions
that get or extract a condition or joins data or function
outputs.](images/function-flow-population.svg)

### Inclusion events

```{r, include=FALSE}
library(dplyr)
library(osdc)
```

#### HbA1c tests above the diagnosis cut-off value (48 mmol/mol or 6.5%)

The function `include_hba1c()` uses `lab_forsker` as the input data to
extract all events of HbA1c tests above the diagnosis cut-off value.

Since the HbA1c diagnosis cut-off value depends on the kind of test that
is used, the inclusion event is defined as follows:

-   For HbA1c IFCC (NPU03835), we include values \>= 6.5 %.
-   For HbA1c DCCT (NPU27300), we include values \>= 48 mmol/mol.

```{r, echo=FALSE}
algorithm |>
  filter(name == "hba1c") |>
  knitr::kable(caption = "Algorithm used in the implementation for including HbA1c.")
```

#### Hospital diagnosis of diabetes

The function `include_diabetes_diagnoses()` uses the hospital contacts
from LPR2 and 3 to include all dates of diabetes diagnoses. Diabetes
diagnoses from both ICD 8 and ICD 10 are included.

This function contains two helper functions:

-   `keep_diabetes_icd10()`
-   `keep_diabetes_icd8()`

<!-- TODO: Add details on how this filtering should be done, e.g., diagnosis codes -->

<!-- TODO: Which specific ICD 8 and 10 codes are included? -->

#### Diabetes-specific podiatrist services

The function `include_podiatrist_services()` uses `sysi` or `sssy` as
input to extract the dates of all diabetes-specific podiatrist services.

<!-- TODO: Add details on how this filtering should be done -->

#### GLD purchases

The function `include_gld_purchases()` uses `lmdb` to extract the dates
of all GLD purchases (from 1997 onwards).

<!-- TODO: Add details on how this filtering should be done -->

<!-- TODO: Add this + link to resource "For details about this, see [link]." -->

### Exclusion events

#### HbA1c tests and GLD purchases during pregnancy

The function `exclude_pregnancy()` uses diagnoses from LPR2 or LPR3 as
input and is used to exclude both HbA1c tests and GLD purchases during
pregnancy.

Internally, this relies on the function `get_pregnancy_dates()` that
contains the following three helper functions:

-   `calculate_pregnancy_index_date_for_mc_visits_wo_end_date()` (this
    might be removed with the inclusion of the birth register)
-   `get_pregnancy_end_dates()`: Keep maternal care visits with an end
    date and drop visits between 40 weeks before end date and 12 weeks
    after end date.
-   `get_maternal_care_visit_dates_without_end_date()`: Uses the output
    from `get_pregnancy_end_dates()` which identifies maternal care
    visits *with* end dates to derive maternal care visits *without* end
    dates. below.

<!-- TODO: What is done with the mc visits without end dates then? -->

<!-- TODO: Add details on how this filtering should be done -->

#### Glucose-lowering brand drugs for weight loss

The function `exclude_wld_purchases()` uses lmdb as input and excludes
the brand drugs Saxenda and Wegovy.

<!-- TODO: Add details on how this filtering should be done -->

#### Metformin purchases for women below age 40

The function `exclude_potential_pcos()` as input to exclude all
purchases of metformin by women below age 40 (i.e., \<= 39 years old) at
the date of purchase. It relies on `bef` as input.

This function contains two helper functions:

-   `keep_women()`
-   `drop_age_40_below()`

<!-- TODO: Add details on how this filtering should be done -->

### Get diagnosis date

The function `get_diagnosis_date()` combines the outputs from the
inclusion and exclusion functions to get the final diagnosis date.
Initially, it drops the first inclusion and exclusion events from the
function outputs with the helper `drop_first_event()`, so that only
those with two or more events are kept. This is then used to assign an
initial diagnosis according to OSDC. Then, all the outputs are joined
together with `join_diagnosis_dates()`.

Finally, the dates outside of the data coverage period are dropped with
`drop_diagnosis_dates_outside_coverage()` to end with a final diagnosis
date. For details on this censoring based on periods with insufficient
data coverage, see the `vignette("design")`.

### Classifying the diabetes type

The next step of the OSDC algorithm classifies individuals from the
extracted diabetes population as having either T1D or T2D. As described
in the `vignette("design")`, individuals not classified as T1D cases are
classified as T2D cases.

The output is a `data.frame` that includes one row per individual in the
diabetes population: one column with their PNR, two columns with
inclusion dates (one "stable" date and one "raw" date - see the
`vignette("design")` for an elaboration on what that entails), and one
column with the diabetes type.

<!-- TODO: add a link to the specific section where this is described -->

![Flow of functions for classifying diabetes status using the `osdc`
package.](images/function-flow-classification.svg)

#### Type 1 classification

The details for the classification of type 1 diabetes is described in
`vignette("design")`. To classify whether an individual has T1D, the
OSDC algorithm includes the following criteria:

1.  `get_t1d_primary_diagnosis()`, which relies on the hospital
    diagnoses extracted from `lpr_diag` (LPR2) and `diagnoser` (LPR3) in
    the previous steps.
2.  `get_only_insulin_purchases()` which relies on the GLD purchases
    from Lægemiddelsdatabasen to get patients where all GLD purchases
    are insulin only.
3.  `get_majority_of_t1d_diagnoses()` (as compared to T2D diagnoses)
    which again relies on primary hospital diagnoses from LPR.
4.  `get_insulin_purchase_within_180_days()` which relies on both
    diagnosis from LPR and GLD purchases from Lægemiddelsdatabasen.
5.  `get_insulin_is_two_thirds_of_gld_doses` which relies on the GLD
    purchases from Lægemiddelsdatabasen.

Note the following hierarchy in first function above: First, the
function checks whether the individual has primary diagnoses from
endocrinological specialty. If that's the case for a given person, the
check of whether they have a majority of T1D primary diagnoses are based
on data from endocrinological specialty. If that's not the case, the
check will be based on primary diagnoses from medical specialties.

#### Type 2 classification

As described in the `vignette("design")`, individuals not classified as
type 1 cases are classified as type 2 cases.
