---
title: "Internal function flow"
---

The function flow describes the functions within the package, both
internal and user-facing, which data sources they rely on, and how they
are connected to each other. First, the functions for classifying
diabetes status are presented, followed by the functions for classifying
the diabetes type.

## Function flow

This results in the functionality flow for classifying diabetes status
seen below. This flow can be divided into two sections: extracting the
diabetes population and classifying diabetes type which we will detail
in the following sections.

![Flow of functions, as well as their required input registers, for
classifying diabetes status using the osdc package. Light blue and
orange boxes represent filtering functions (inclusion and exclusion
events, respectively). Uncoloured boxes are helper functions that get or
extract a condition or joins data or function
outputs.](function-flow.svg)

## Population extraction

In the following sections, we describe the functions used to extract the
diabetes population from the Danish registers. The functions are divided
into inclusion and exclusion events, and the final diagnosis date is
calculated based on these events.

![Flow of functions, as well as their required input registers, for
extracting the population with diabetes using the osdc package. Light
blue and orange boxes represent filtering functions (inclusion and
exclusion events, respectively). Uncoloured boxes are helper functions
that get or extract a condition or joins data or function
outputs.](function-flow-population.svg)

## Inclusion events

```{r, include=FALSE}
library(dplyr)
library(osdc)
```

### `join_lpr2()`

```{r}
#' Process and join the two LPR2 registers to extract diabetes diagnoses data.
#'
#' The output is used as inputs to `include_diabetes_diagnoses()` and to
#' `get_pregnancy_dates()` (see exclusion events).
#'
#' @param lpr_diag The LPR2 register containing diabetes diagnoses.
#' @param lpr_adm The LPR2 register containing hospital admissions.
#'
#' @return The same type as the input data, default as a [tibble::tibble()],
#'  with the following columns:
#'
#'  -   `pnr`: The personal identification variable.
#'  -   `date`: The date of all the recorded diagnosis (renamed from `d_inddto`).
#'  -   `is_primary_diagnosis`: Whether the diagnosis was a primary diagnosis.
#'  -   `has_t1d`: Whether the diagnosis was T1D-specific.
#'  -   `has_t2d`: Whether the diagnosis was T2D-specific.
#'  -   `has_pregnancy_event`: Whether the person has an event related to pregnancy like giving birth or having a miscarriage at the given date.
#'  -   `is_endocrinology_department`: Whether the diagnosis was made made by an
#'      endocrinology (TRUE) or other medical (FALSE) department.
#'
#' @keywords internal
#' @inherit algorithm seealso
#'
#' @examples
#' sim_data <- simulate_registers(c("lpr_diag", "lpr_adm"), 100)
#' join_lpr2(
#'   lpr_diag = sim_data$lpr_diag,
#'   lpr_adm = sim_data$lpr_adm
#' )
join_lpr2 <- function(lpr_diag, lpr_adm) {
  # Filter using the algorithm for LPR2
  lpr_diag |>
    dplyr::full_join(lpr_adm, by = dplyr::join_by(.data$recnum)) |>
    dplyr::select(
      pnr,
      date = d_inddto
      # is_primary_diagnosis =
      # has_t1d =
      # has_t2d =
      # has_pregnancy_event =
      # is_endocrinology_department =
    )
}
```

### `join_lpr3()`

```{r}
#' Process and join the two LPR3 registers to extract diabetes diagnoses data.
#'
#' The output is used as inputs to `include_diabetes_diagnoses()` and to
#' `get_pregnancy_dates()` (see exclusion events).
#'
#' @param diagnoser The LPR3 register containing diabetes diagnoses.
#' @param kontakter The LPR3 register containing hospital contacts/admissions.
#'
#' @return The same type as the input data, default as a [tibble::tibble()],
#'  with the following columns:
#'
#'  -   `pnr`: The personal identification variable.
#'  -   `date`: The date of all the recorded diagnosis (renamed from `d_inddto`).
#'  -   `is_primary_diagnosis`: Whether the diagnosis was a primary
#'      diagnosis.
#'  -   `has_t1d`: Whether the diagnosis was T1D-specific
#'  -   `has_t2d`: Whether the diagnosis was T2D-specific.
#'  -   `has_pregnancy_event`: Whether the person has an event related to pregnancy like giving birth or having a miscarriage at the given date.
#'  -   `is_endocrinology_department`: Whether the diagnosis was made made by an
#'      endocrinology (TRUE) or other medical (FALSE) department.
#'
#' @keywords internal
#' @inherit algorithm seealso
#'
#' @examples
#' sim_data <- simulate_registers(c("diagnoser", "kontakter"), 100)
#' join_lpr3(
#'   diagnoser = sim_data$diagnoser,
#'   kontakter = sim_data$kontakter
#' )
join_lpr3 <- function(diagnoser, kontakter) {
  # Filter using the algorithm for LPR3
  diagnoser |>
    dplyr::full_join(kontakter, by = dplyr::join_by(.data$dw_ek_kontakt)) |>
    # Ensure the values are always lower case.
    dplyr::mutate(hovedspeciale_ans = tolower(hovedspeciale_ans)) |>
    dplyr::select(
      "pnr" = "cpr",
      "date" = "dato_start"
      # is_primary_diagnosis =
      # has_t1d =
      # has_t2d =
      # has_pregnancy_event =
      # is_endocrinology_department =
    )
}
```

### `include_diabetes_diagnosis()`

```{r}
#' Include diabetes diagnoses from LPR2 and LPR3.
#'
#' Uses the hospital contacts from LPR2 and LPR3 to include all dates of diabetes
#' diagnoses to use for inclusion, as well as additional information needed to classify diabetes
#' type. Diabetes diagnoses from both ICD-8 and ICD-10 are included.
#'
#' The output is used as inputs to `join_inclusions()`.
#' This output is passed to the `join_inclusions()` function, where the
#' `dates` variable is used for the final step of the inclusion process.
#' The variables of counts of diabetes type-specific primary diagnoses (the
#' four columns prefixed `n_` above) are carried over for the subsequent
#' classification of diabetes type, initially as inputs to the
#' `get_t1d_primary_diagnosis()` and `get_majority_of_t1d_diagnoses()`
#' functions.
#'
#' @param lpr2 The output from `join_lpr2()`.
#' @param lpr3 The output from `join_lpr3()`.
#'
#' @return The same type as the input data, default as a [tibble::tibble()],
#'  with the following columns and up to two rows per individual:
#'
#'  -   `pnr`: The personal identification variable.
#'  -   `dates`: The dates of the first and second hospital diabetes diagnosis.
#'  -   `n_t1d_endocrinology`: The number of type 1 diabetes-specific primary
#'      diagnosis codes from endocrinology departments.
#'  -   `n_t2d_endocrinology`: The number of type 2 diabetes-specific primary
#'      diagnosis codes from endocrinology departments.
#'  -   `n_t1d_medical`: The number of type 1 diabetes-specific primary
#'      diagnosis codes from medical departments.
#'  -  `n_t2d_medical`: The number of type 2 diabetes-specific primary
#'      diagnosis codes from medical departments.
#'  -  `has_lpr_diabetes_diagnosis`: A logical variable that acts as a helper
#'      indicator for use in later functions.
#'
#' @keywords internal
#' @inherit algorithm seealso
#'
#' @examples
#' sim_data <- simulate_registers(c("lpr_diag", "lpr_adm", "diagnoser", "kontakter"), 100)
#' include_diabetes_diagnosis(
#'   lpr2 = join_lpr2(sim_data$lpr_diag, sim_data$lpr_adm),
#'   lpr3 = join_lpr3(sim_data$diagnoser, sim_data$kontakter)
#' )
include_diabetes_diagnosis <- function(lpr2, lpr3) {
  # Combine and process the two inputs
  lpr2 |>
    dplyr::full_join(lpr3, by = dplyr::join_by(.data$pnr)) |>
    dplyr::select(
      "pnr",
      "dates" = "date"
      # n_t1d_endocrinology =
      # n_t2d_endocrinology =
      # n_t1d_medical =
      # n_t2d_medical =
    ) |>
    dplyr::mutate(has_lpr_diabetes_diagnosis = TRUE)
}
```

### `include_podiatrist_services()`

See `?include_podiatrist_services` for more information.

### `include_hba1c()`

See `?include_hba1c` for more information.

### `include_gld_purchases()`

See `?include_gld_purchases` for more information.

## Exclusion events

### `exclude_potential_pcos()`

See `?exclude_potential_pcos` for more information.

### `exclude_pregnancy()`

```{r}
#' Exclude any pregnancy events that could be gestational diabetes.
#'
#'
#' The function `exclude_pregnancy()` takes the combined outputs from
#' `join_lpr2()`, `join_lpr3()`, `include_hba1c()`, and
#' `exclude_potential_pcos()` and uses diagnoses from LPR2 or LPR3 to
#' exclude both elevated HbA1c tests and GLD purchases during pregnancy, as
#' these may be due to gestational diabetes, rather than type 1 or type 2
#' diabetes. The aim is to identify pregnancies based on diagnosis codes
#' specific to pregnancy-ending events (e.g. live births or miscarriages),
#' and then use the dates of these events to remove inclusion events in the
#' preceding months that may be related to gestational diabetes (e.g.
#' elevated HbA1c tests or purchases of glucose-lowering drugs during
#' pregnancy).
#'
#' After these exclusion functions have been applied, the output serves as
#' inputs to two sets of functions:
#'
#' 1.  The censored HbA1c and GLD data are passed to the
#'     `join_inclusions()` function for the final step of the inclusion
#'     process.
#' 2.  the censored GLD data is passed to the
#'     `get_only_insulin_purchases()`,
#'     `get_insulin_purchases_within_180_days()`, and
#'     `get_insulin_is_two_thirds_of_gld_doses()` helper functions for the
#'     classification of diabetes type.
#'
#' @param excluded_pcos Ouptut from `exclude_potential_pcos()`.
#' @param pregnancy_dates Output from `get_pregnancy_dates()`.
#' @param included_hba1c Output from `include_hba1c()`.
#'
#' @returns The same type as the input data, default as a [tibble::tibble()].
#'    Has the same output data as the input `excluded_potential_pcos()`, except
#'    for a helper logical variable `no_pregnancy` that is used in later functions.
#' @keywords internal
#' @inherit algorithm seealso
#'
#' @examples
#' sim_data <- simulate_registers(c(
#'   "lmdb", "bef", "lpr_diag", "lpr_adm",
#'   "diagnoser", "kontakter", "lab_forsker"
#' ), 100)
#' lpr2 <- join_lpr2(sim_data$lpr_diag, sim_data$lpr_adm)
#' lpr3 <- join_lpr3(sim_data$diagnoser, sim_data$kontakter)
#' lmdb |>
#'   include_gld_purchases() |>
#'   exclude_potential_pcos(sim_data$bef) |>
#'   exclude_pregnancy(
#'     get_pregnancy_dates(lpr2, lpr3),
#'     include_hba1c(sim_data$lab_forsker)
#'   )
exclude_pregnancy <- function(excluded_pcos, pregnancy_dates, included_hba1c) {
  # Filter using the algorithm for pregnancy
  excluded_pcos |>
    # Exclude those who are not pregnant.
    dplyr::full_join(pregnancy_dates, by = dplyr::join_by(.data$pnr)) |>
    dplyr::full_join(included_hba1c, by = dplyr::join_by(.data$pnr)) |>
    # Filtering here...
    dplyr::mutate(no_pregnancy = TRUE)
}
```

```{r}
#' Simple function to get only the pregnancy event dates.
#'
#' @param lpr2 Output from `join_lpr2()`.
#' @param lpr3 Output from `join_lpr3()`.
#'
#' @returns The same type as the input data, default as a [tibble::tibble()].
#' @keywords internal
#' @inherit algorithm seealso
#'
#' @examples
#' sim_data <- simulate_registers(c("lpr_diag", "lpr_adm", "diagnoser", "kontakter"), 100)
#' lpr2 <- join_lpr2(sim_data$lpr_diag, sim_data$lpr_adm)
#' lpr3 <- join_lpr3(sim_data$diagnoser, sim_data$kontakter)
#' get_pregnancy_dates(lpr2, lpr3)
get_pregnancy_dates <- function(lpr2, lpr3) {
  # Filter using the algorithm for pregnancy
  lpr2 |>
    dplyr::full_join(lpr3, by = dplyr::join_by(pnr)) |>
    dplyr::filter(has_pregnancy_events) |>
    dplyr::select(
      pnr,
      pregnancy_event_date = date
    )
}
```

## Joining inclusions and exclusions

### `join_inclusions()`

The function `join_inclusions()` appends/row-binds the dates output from
functions the process the four types of inclusion events by `pnr`. Thus,
it takes as input the following variables output from the following
functions:

-   From `include_diabetes_diagnoses()`:
    -   `pnr`: identifier variable
    -   `dates`: dates of the first and second hospital diabetes
        diagnosis
-   From `include_podiatrist_services()`
    -   `pnr`: identifier variable
    -   `dates`: the dates of the first and second diabetes-specific
        podiatrist record
-   From `exclude_pregnancy()`:
    -   `pnr`: identifier variable
    -   `dates`: the dates of the first and second elevated HbA1c test
        results (after censoring)
-   From `exclude_pregnancy()`:
    -   `pnr`: identifier variable
    -   `date`: dates of all purchases of GLD
        -   The dates of the first and second purchase of GLD of each
            individual are extracted from these and appended as two rows
            to the ´dates´ variable.

The output from the function is a `data.frame` containing two variables
(`pnr` and `dates`) and 1 to 8 rows per ´pnr´. This output is passed to
`get_inclusion_date()`.

### Get diagnosis date

```{r}
#' Get the final diagnosis date based on all the inclusion event types.
#'
#' The function `get_inclusion_date()` takes the output from `join_inclusions()`
#' and defines the final diagnosis date based on all the inclusion event types.
#' Keeps only those with 2 or more recorded inclusion events, regardless of the
#' type of these events (e.g. two elevated HbA1c tests will lead to inclusion as
#' well as one elevated HbA1c test followed by a purchase of glucose-lowering
#' drugs).
#'
#' @param inclusions Output from [join_inclusions()].
#' @param stable_inclusion_start_date The date from when the inclusion events
#'    from all sources are considered more 'stable' (e.g. time after the change
#'    in how medication drugs are labeled and how doctors actually regularly
#'    input the new change into the database).
#'
#' @returns The same type as the input data, default as a [tibble::tibble()],
#'   with the following columns:
#'
#'   - `pnr`: The person identifier variable.
#'   - `raw_inclusion_date`: Date of inclusion, which is the second
#'      earliest recorded event.
#'   - `stable_inclusion_date`: Date of inclusion of individuals included
#'      at least one year after the incorporation of inclusions based on
#'      glucose-lowering drug data (1998 onwards when using National Patient
#'      Register data for censoring of gestational diabetes). Limits the
#'      included cohort to only individuals with a valid date of inclusion
#'      (and thereby valid age at inclusion & duration of diabetes).
#'
#' @keywords internal
#' @inherit algorithm seealso
get_inclusion_date <- function(inclusions, stable_inclusion_start_date = "1998-01-01") {
  inclusions |>
    # TODO: May need to consider more efficient ways than using group by.
    dplyr::group_by(.data$pnr) |>
    # Drop earliest date so only those with two or more events are included.
    dplyr::filter(.data$dates != min(.data$dates, na.rm = TRUE)) |>
    dplyr::mutate(
      # Earliest date in the rows for each individual.
      raw_inclusion_date = min(.data$dates, na.rm = TRUE),
      stable_inclusion_date = dplyr::if_else(
        .data$raw_inclusion_date < lubridate::as_date(stable_inclusion_start_date),
        NA,
        .data$raw_inclusion_date
      )
    ) |>
    dplyr::select(
      .data$pnr,
      .data$raw_inclusion_date,
      .data$stable_inclusion_date
    )
}
```

### Classifying the diabetes type

The next step of the OSDC algorithm classifies individuals from the
extracted diabetes population as having either T1D or T2D. As described
in the `vignette("design")`, individuals not classified as T1D cases are
classified as T2D cases.

As the diabetes type classification incorporates an evaluation of the
time from diagnosis/inclusion to first subsequent purchase of insulin,
the `get_diabetes_type()` function has to take the date of diagnosis and
all purchases of GLD drugs (after censoring) as inputs. In addition,
information on diabetes type-specific primary diagnoses from hospitals
is also a requirement.

Thus, the function takes the following inputs from
`get_inclusion_date()`, `exclude_pregnancy()`, and
`include_diabetes_diagnoses()`:

-   From `get_inclusion_date()`: Information on date of diagnosis of
    diabetes
    -   `pnr`
    -   `raw_inclusion_date`
    -   `stable_inclusion_date`
-   From `exclude_pregnancy()`: Information on historic GLD purchases:
    -   `pnr`: identifier variable
    -   `date`: dates of all purchases of GLD.
    -   `atc`: type of drug
    -   `contained_doses`: defined daily doses of drug contained in
        purchase
-   From `include_diabetes_diagnoses()`: Information on diabetes
    type-specific primary diagnoses from hospitals:
    -   `pnr`: identifier variable
    -   `n_t1d_endocrinology`: number of type 1 diabetes-specific
        primary diagnosis codes from endocrinological departments
    -   `n_t2d_endocrinology`: number of type 2 diabetes-specific
        primary diagnosis codes from endocrinological departments
    -   `n_t1d_medical`: number of type 1 diabetes-specific primary
        diagnosis codes from medical departments
    -   `n_t2d_medical`: number of type 2 diabetes-specific primary
        diagnosis codes from medical departments

For each `pnr` number, several helper functions are applied to these
inputs to extract additional information from the censored GLD data and
diagnoses to use for classification of diabetes type. All of these
return a single value (`TRUE`, otherwise `FALSE`) for each individual:

-   `get_only_insulin_purchases()`:
    -   Inputs passed from `exclude_pregnancy()`:
        -   `atc`
    -   Outputs:
        -   only_insulin_purchases = `TRUE` if no purchases with `atc`
            starting with "A10A" are present
-   `get_insulin_purchases_within_180_days()`
    -   Inputs passed from `exclude_pregnancy()`:
        -   `date` & `atc`
    -   Inputs passed from `get_inclusion_date()`:
        -   `raw_inclusion_date`
    -   Outputs: `TRUE` If any purchases with `atc` starting with "A10A"
        have a `date` between 0 and 180 days higher than
        `raw_inclusion_date`
-   `get_insulin_is_two_thirds_of_gld_doses()`
    -   Inputs passed from `exclude_pregnancy()`:
        -   `contained_doses` & `atc`
    -   Outputs: `TRUE` If the sum of `contained_doses` of rows of `atc`
        starting with "A10A" (except "A10AE5") is at least twice the sum
        of `contained_doses` of rows of `atc` starting with "A10B" or
        "A10AE5"
-   `get_any_t1d_primary_diagnoses()`:
    -   Inputs passed from `include_diabetes_diagnoses()`:
        -   `n_t1d_endocrinology` & `n_t1d_medical`
    -   Outputs: `TRUE` if the combined sum of the inputs is 1 or above.
-   `get_type_diagnoses_from_endocrinology()`:
    -   Inputs passed from `include_diabetes_diagnoses()`:
        -   `n_t1d_endocrinology`, `n_t2d_endocrinology`
    -   Outputs: `type_diagnoses_from_endocrinology` = `TRUE` if the
        combined sum of the inputs is 1 or above
-   `get_type_diagnosis_majority()`:
    -   Inputs passed from `include_diabetes_diagnoses()`:
        -   `n_t1d_endocrinology`, `n_t2d_endocrinology`,
            `n_t1d_medical` & `n_t2d_medical`
    -   Inputs passed from `get_type_diagnoses_from_endocrinology()`:
        -   `type_diagnoses_from_endocrinology`
    -   Outputs: `TRUE` if `type_diagnoses_from_endocrinology` == `TRUE`
        and `n_t1d_endocrinology` is above `n_t2d_endocrinology`. Also
        `TRUE` if `type_diagnoses_from_endocrinology` = `FALSE` and
        `n_t1d_medical` is above `n_t2d_medical`

`get_diabetes_type()` evaluates all the outputs from the helper
functions to define diabetes type for each individual. Diabetes type is
classified as "T1D" if:

-   `only_insulin_purchases` == `TRUE` & `any_t1d_primary_diagnoses` ==
    `TRUE`
-   Or `only_insulin_purchases` == `FALSE` & `any_t1d_primary_diagnoses`
    == `TRUE` & `type_diagnosis_majority` == `TRUE` &
    `insulin_is_two_thirds_of_gld_doses` == `TRUE` &
    `insulin_purchases_within_180_days` == `TRUE`

`get_diabetes_type()` returns a `data.frame` with one row per `pnr`
number and four columns: `pnr`, `stable_inclusion_date`,
`raw_inclusion_date` & `diabetes_type`. This is the final product of the
OSDC algorithm. See the `vignette("design")` for an more detail on the
two inclusion dates and their intended use-cases.

<!-- TODO: Create updated image similar to https://aastedet.github.io/dissertation/4-results.html#fig-osdc-type-flow to reflect the new diabetes type logic and embed image here for reference-->

<!-- TODO:  The following explanatory sections on T1D and T2D classification need to be aligned with the technical sections above, and possibly moved up to them-->

![Flow of functions for classifying diabetes status using the `osdc`
package.](function-flow-classification.svg)

#### Type 1 classification

The details for the classification of type 1 diabetes is described in
`vignette("design")`. To classify whether an individual has T1D, the
OSDC algorithm includes the following criteria:

1.  `get_t1d_primary_diagnosis()`, which relies on the hospital
    diagnoses extracted from `lpr_diag` (LPR2) and `diagnoser` (LPR3) in
    the previous steps.
2.  `get_only_insulin_purchases()` which relies on the GLD purchases
    from Lægemiddeldatabasen to get patients where all GLD purchases are
    insulin only.
3.  `get_majority_of_t1d_diagnoses()` (as compared to T2D diagnoses)
    which again relies on primary hospital diagnoses from LPR.
4.  `get_insulin_purchase_within_180_days()` which relies on both
    diagnosis from LPR and GLD purchases from Lægemiddeldatabasen.
5.  `get_insulin_is_two_thirds_of_gld_doses` which relies on the GLD
    purchases from Lægemiddeldatabasen.

Note the following hierarchy in first function above: First, the
function checks whether the individual has primary diagnoses from
endocrinological specialty. If that's the case for a given person, the
check of whether they have a majority of T1D primary diagnoses are based
on data from endocrinological specialty. If that's not the case, the
check will be based on primary diagnoses from medical specialties.

#### Type 2 classification

As described in the `vignette("design")`, individuals not classified as
type 1 cases are classified as type 2 cases.
